Comparing variations in sediment transport capacity between subglacial and subaerial systems

### 2023-03-21 theory.tex
small change

  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to subaerial ones (Figure~\ref{fig:range}).
  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S \ref{fig:gammas}).

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is greater than $1.5$.
  These results show that increased variability occurs across a wide range of subaerial and subglacial channel configurations (Figure~\ref{fig:range} g, h).
The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, a smaller amount of relative variability compared to  water velocity $\frac{\chi(v_{g})}{\chi(v_{f})}$ or shear stress $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$ (Figure~\ref{fig:range}).
Consistent with the individual case (Figure ~\ref{fig:model_outs}), the subaerial systems' reduce variability in width-integrated shear stress occurs  as the evolving channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}). 
\ian{extra sentence is needed here}

@@

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
  Yet, reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport and the tendency for glaciers to transport sediment in a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
  Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.

  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
  Increased sediment discharge may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will cause a large increase in sediment transport capacity (Section~\ref{sect:sub_mode}).

@@

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
  The rapid increase and decreases in sediment transport capacity over a diurnal cycle in a subglacial system may compound the variability already present in subaerial systems \citep{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
  Aside from the sporadic nature of these events, the observed flushings could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.

>>>>

  Across the accepted model runs, variability in water velocity and shear stress remains substantially higher across the subglacial model compared to subaerial ones (Figure~\ref{fig:range}).
  Water velocity, shear stress, and width-integrated shear stress all have a variability ratio ($\frac{\chi(\Gamma_{g})}{\chi(\Gamma_f)}$) greater than $1$ in the subglacial  system compared to subaerial one (Figure~S \ref{fig:gammas}).
  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range}\,d--f and S \ref{fig:gammas}).

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is greater than $1.5$.
  These results show that increased variability occurs across a wide range of subaerial and subglacial channel configurations (Figure~\ref{fig:range} g, h).
The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$, a smaller amount of relative variability compared to  water velocity $\frac{\chi(v_{g})}{\chi(v_{f})}$ or shear stress $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$ (Figure~\ref{fig:range}).
Consistent with the individual case (Figure ~\ref{fig:model_outs}), the subaerial systems' reduce variability in width-integrated shear stress occurs  as the evolving channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
  Importantly though, no runs experienced relative variability in width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$, showing that the subglacial has consistently high variability in width integrated shear stress compared to subaerial systems across the range of tested situations.

@@

  The response of sediment discharge caused by the greater variability in sediment transport in subglacial channels will be greatest in completely transport-limited regimes \citep[e.g.][]{kasmalkar2019}.
  In supply-limited systems, the changing sediment transport capacity (Figure~\ref{fig:model_outs}) would have a minimal effect on sediment discharge due to the absence of sediment available for transport \citep{delaney2019}.
  Yet, in the subglacial system reaching the threshold of motion for sediment frequently, across a range of discharges,  may result in additional sediment transport.
  The variable sediment transport capacity may result in a tendency for glaciers to evacuate sediment and transition to a supply-limited regime whereby sediment discharge scales with glacier sliding  \citep[e.g.][]{herman2015,koppes2015}.
  Sediment exhaustion by this process may also explain the dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, as opposed to glacier melt \citep{overeem2017}.

  The effect of pressurized channel flow could be most pronounced during precipitation or flood events when subglacial conduits have not adapted to the water discharge flowing through them, resulting in a large increase in sediment transport capacity and discharge \citep[e.g.][]{cowan1988,delaney2019}.
  Exceptional sediment discharge during floods may result from changes to sediment access below the glacier, yet rapid increases in water velocity from water flowing through a small channel will also cause a large increase in sediment transport capacity, not possible in a subaerial system (Section~\ref{sect:sub_mode}).

@@

  Large variations in sediment transport capacity below glaciers may result in sediment mobilization and deposition processes in close spatial and temporal proximity to each other below the glacier \citep{gimbert2016,perolo2018}.
  The rapid increase and decreases in sediment transport capacity, for example over a diurnal cycle, in a subglacial system may compound the variability in sediment transport already present in subaerial systems \citep{williams1989,jerolmack2010}.
  As a result, variations in sediment transport capacity could be responsible for the fluctuations in sediment discharge (``flushings'') from glaciers that are not attributed to variations in water discharge \citep[e.g.][]{richards2003,swift2021}.
  Aside from the sporadic nature of these events, the observed flushings could result from increases in water velocity resulting from reduced channel size, not water discharge, which increase sediment transport capacity.
